---
title: "p8105_hw3_ep2899"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup}
library(tidyverse)
library(p8105.datasets)
```

###Problem 0


### Problem 1
```{r}
data("instacart")
```

This dataset contains 'r nrow(instacart)' rows and ... columns

Observations are the level of items in orders by user. There are user/order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and same numeric codes.

how many aisles, and which are most items from?

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

Let's make a plot

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle, n)
  ) %>% 
  ggplot(aes (x = aisle, y=n)) +
  geom_point() +
  theme(axis.text.x = element_text(angle =90, vjust = 0.5, hjust =1))
```

Let's make a table!

```{r}
instacart %>% 
  filter (aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits") ) %>% 
    group_by(aisle) %>% 
count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter (rank <4) %>% 
  arrange(aisle, rank) %>% 
  knitr::kable()
```

Apples vs ice cream..

```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day))
```

